---
title: "Personality Dynamics - Power Calculation"
author: Clemens Stachl
output: html_notebook
---

As mentioned in the text, false positive results want to be avoided. 
Therefore the number of planned hypothesis tests should be accounted for in the 
power analysis. How many significance tests will you perform?

# 1. Adjust for multiple comparisons
```{r}
# number of hypothesis tests
r <- 2

# your significance level (family-wise) e.g. 0.05
pi <- 0.05

# calculate your corrected alpha
alpha_corr <- pi/r

# use alpha_corr in the power calculations e.g. if you want to calculate 2 correlations
# simply put alpha_corr into your power calculation(s) instead of 0.05

```

# Power Calculations     
```{r, echo=TRUE}
library(pwr)

# read also
#https://cran.r-project.org/web/packages/pwr/vignettes/pwr-vignette.html

```

## T-tests
```{r}
# Conventions:	s 0.2	m 0.5	l 0.8
?pwr.t.test
pwr.t.test(d=0.25, sig.level = alpha_corr, power = 0.8, alternative = "two.sided", type = "two.sample")

```


## ANOVA
```{r}
# Conventions: s 0.1	m 0.25	l 0.4
pwr.anova.test(k = 3, sig.level = alpha_corr, power = 0.8, f = 0.25)

```

## F-Test (for regression) - single and multiple regression
Please note that this test is rather uninformative (correlations between predictors are not accounted for), as a supplement use the correlations power analysis below (for two-variable linear associations).

- n: sample size     
- u: df numerator degrees of freedom (number of predictors without the intercept)      
- v: df denominator degress of freedom (v = n - u - 1)  therefore n = v + u + 1  
- f2: effect size R^2/(1-R^2)     

```{r}
# Conventions: s 0.02 m 0.15 l 0.35
fpower <- pwr.f2.test(u = 5, f2 = 0.10/(1-0.10), sig.level = alpha_corr, power = 0.8)
# min n needed
ceiling(fpower$v) + fpower$u + 1

```

## Correlation
```{r}

#Conventions: s 0.1	m 0.3	l 0.5
pwr.r.test(r = .21, sig.level = alpha_corr, power = 0.8, alternative = "two.sided")

```

## Multilevel Models

<!-- http://rpsychologist.com/introducing-powerlmm -->


```{r}
install.packages("powerlmm")
library(powerlmm)

# tutorial from: http://rpsychologist.com/introducing-powerlmm 

# dropout per treatment group
d <- per_treatment(control = dropout_weibull(0.3, 2),
              treatment = dropout_weibull(0.2, 2))

# Setup design
p <- study_parameters(n1 = 11, # time points
                      n2 = 10, # subjects per cluster
                      n3 = 5, # clusters per treatment arm
                      icc_pre_subject = 0.5,
                      icc_pre_cluster = 0,
                      icc_slope = 0.05,
                      var_ratio = 0.02,
                      dropout = d,
                      cohend = -0.8)

# Power
get_power(p)





```




